Supplemental Material: Multidimensional analysis and detection of informative features in human white matter

### 2021-01-21 reports/manuscript/abstract.tex
ENH: update paper and figures

\section*{Abstract}

The white matter contains long-range connections between different
brain regions and the organization of these connections holds important
implications for brain function in health and disease. Tractometry
uses diffusion-weighted magnetic resonance imaging (dMRI) data
to quantify tissue properties (e.g. fractional anisotropy (FA),
mean diffusivity (MD), etc.), along the trajectories of these
connections\cite{yeatman2012tract}. Statistical inference from
tractometry usually either (a) averages these quantities along
the length of each bundle in each individual, or (b) performs analysis
point-by-point along each bundle, with group comparisons or regression
models computed separately for each point along every one of the bundles.
These approaches are limited in their sensitivity, in the former case, or
in their statistical power, in the latter.
In the present work, we developed a method based on the sparse group
lasso (SGL) \cite{simon2013sparse} that takes into account tissue
properties measured along all of the bundles, and selects informative
features by enforcing sparsity, not only at the level of individual
bundles, but also across the entire set of bundles and all of the measured
tissue properties. The sparsity penalties for each of these constraints
is identified using a nested cross-validation scheme that guards
against over-fitting and simultaneously identifies the correct
level of sparsity. We demonstrate the accuracy of the method in two
settings: i) In a classification setting, patients with amyotrophic
lateral sclerosis (ALS) are accurately distinguished from matched
controls \cite{sarica2017corticospinal}. Furthermore, SGL automatically
identifies FA in the corticospinal tract as important for this
classification -- correctly finding the parts of the white matter known
to be affected by the disease. ii) In a regression setting, dMRI is
used to accurately predict ``brain age'' \cite{yeatman2014lifespan,
Brown2012-so}. In this case, the weights are distributed throughout the
white matter indicating that many different regions of the white matter
change with development and contribute to the prediction of age. Thus,
SGL makes it possible to leverage the multivariate relationship between
diffusion properties measured along multiple bundles to make accurate
predictions of subject characteristics while simultaneously discovering
the most relevant features of the white matter for the characteristic of
interest.

>>>>

\begin{abstract}
    The white matter contains long-range connections between different
    brain regions and the organization of these connections holds important
    implications for brain function in health and disease. Tractometry
    uses diffusion-weighted magnetic resonance imaging (dMRI) data
    to quantify tissue properties (e.g. fractional anisotropy (FA),
    mean diffusivity (MD), etc.), along the trajectories of these
    connections\cite{yeatman2012tract}. Statistical inference from
    tractometry usually either (a) averages these quantities along
    the length of each bundle in each individual, or (b) performs analysis
    point-by-point along each bundle, with group comparisons or regression
    models computed separately for each point along every one of the bundles.
    These approaches are limited in their sensitivity, in the former case, or
    in their statistical power, in the latter.
    In the present work, we developed a method based on the sparse group
    lasso (SGL) \cite{simon2013sparse} that takes into account tissue
    properties measured along all of the bundles, and selects informative
    features by enforcing sparsity, not only at the level of individual
    bundles, but also across the entire set of bundles and all of the measured
    tissue properties. The sparsity penalties for each of these constraints
    is identified using a nested cross-validation scheme that guards
    against over-fitting and simultaneously identifies the correct
    level of sparsity. We demonstrate the accuracy of the method in two
    settings: i) In a classification setting, patients with amyotrophic
    lateral sclerosis (ALS) are accurately distinguished from matched
    controls \cite{sarica2017corticospinal}. Furthermore, SGL automatically
    identifies FA in the corticospinal tract as important for this
    classification -- correctly finding the parts of the white matter known
    to be affected by the disease. ii) In a regression setting, dMRI is
    used to accurately predict ``brain age'' \cite{yeatman2014lifespan,
    Brown2012-so}. In this case, the weights are distributed throughout the
    white matter indicating that many different regions of the white matter
    change with development and contribute to the prediction of age. Thus,
    SGL makes it possible to leverage the multivariate relationship between
    diffusion properties measured along multiple bundles to make accurate
    predictions of subject characteristics while simultaneously discovering
    the most relevant features of the white matter for the characteristic of
    interest.
\end{abstract}

### 2021-01-21 reports/manuscript/paper.tex
Add details to cross-validation section and software implementation. Update some references and minor edits to other sections
ENH: update paper and figures

\documentclass[11pt,twoside,a4paper]{article}
\usepackage{authblk}  % For author affiliations with the basic article class
\usepackage{booktabs}  % For formal tables
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{mathtools}  % for \shortintertext{} and PairedDelimiter
\usepackage{xcolor}

\graphicspath{ {../figures/} }

\DeclarePairedDelimiter\p{(}{)}
\DeclarePairedDelimiter\brkt{[}{]}

\newcommand{\ariel}[1]{{\color{blue} Ariel: [{#1}]}}
\newcommand{\adam}[1]{{\color{green} Adam: [{#1}]}}
\newcommand{\strike}[1]{{\color{red} Strikeout: [{#1}]}}
\newcommand{\authoremail}[1]{\href{mailto:{#1}}{\textnormal{\texttt{{#1}}}}}

% Set the affiliation font and the affiliation separation
\renewcommand\Affilfont{\itshape\small}
\setlength{\affilsep}{1em}
\setcounter{Maxaffil}{3}

@@

\title{AFQ-Insight}
\title{Multidimensional analysis and detection of informative features in diffusion MRI measurements of human white matter}
\date{\today}

\author[1]{Adam Richie-Halford\thanks{Corresponding author: \authoremail{richford@uw.edu}}}
\author[2]{Jason Yeatman}
\author[3]{Noah Simon}
\author[4]{Ariel Rokem}

\affil[1]{University of Washington Department of Physics, Seattle, WA 98105}
\affil[2]{University of Washington Department of Speech and Hearing Sciences, Seattle, WA 98105}
\affil[3]{University of Washington Department of Biostatistics, Seattle, WA 98105}
\affil[4]{University of Washington e{S}cience Institute, Seattle, WA 98105}

\input{abstract}

\maketitle
\tableofcontents

\input{sections}

\bibliographystyle{abbrv}
\bibliography{paper}

>>>>

%% ****** Start of file apstemplate.tex ****** %
%%
%%   This file is part of the APS files in the REVTeX 4 distribution.
%%   Version 4.1r of REVTeX, August 2010
%%
%%
%%   Copyright (c) 2001, 2009, 2010 The American Physical Society.
%%
%%   See the REVTeX 4 README file for restrictions and more information.
%%
%
% This is a template for producing manuscripts for use with REVTEX 4.0
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.
%
% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
%  Add 'draft' option to mark overfull boxes with black boxes
%  Add 'showpacs' option to make PACS codes appear
%  Add 'showkeys' option to make keywords appear
\documentclass[10pt,%
               aps,%
               prl,%
               preprint,%
               superscriptaddress,%
               preprintnumbers,%
               amsmath,%
               floatfix,%
               endfloats*]{revtex4-2}
\usepackage[T1]{fontenc}
\usepackage[utf8x]{inputenc} 
\usepackage[export]{adjustbox}  % Used to adjust figure frames
\usepackage{subcaption}
\usepackage{url}
\usepackage{color}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{float}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{dcolumn}
\usepackage{wrapfig}
\usepackage{physics}
\usepackage[inline]{enumitem}
\usepackage[per-mode=symbol,group-separator={,},group-minimum-digits=4]{siunitx}
\usepackage{todonotes}
\usepackage{graphicx}
\graphicspath{ {../paper_figures} }

%%% HELPER CODE FOR DEALING WITH EXTERNAL REFERENCES
\usepackage{xr-hyper}
\usepackage[pdftex,%
            colorlinks=true,%
            linkcolor=blue,%
            citecolor=blue,%
            urlcolor=blue,%
            bookmarksnumbered=true,%
            bookmarksopen=true]{hyperref}
\usepackage[capitalize]{cleveref}
\makeatletter
\newcommand*{\addFileDependency}[1]{
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[2][ext:]{
    \externaldocument[#1]{#2}
    \addFileDependency{#2.tex}
    \addFileDependency{#2.aux}
}
%%% END HELPER CODE

% put all the external documents here!
\myexternaldocument[supp:]{supplemental}

@@

\title{Multidimensional analysis and detection of informative features in human white matter}

\author{Adam \surname{Richie-Halford}}%
\email{richford@uw.edu}%
\affiliation{%
    eScience Institute,%
    University of Washington, Seattle,%
    Washington 98195--1560, USA
}

\author{Jason \surname{Yeatman}}%
\affiliation{%
    Graduate School of Education and Division of Developmental and Behavioral Pediatrics,%
    Stanford University,%
    Stanford, CA, 94305, USA
}

\author{Noah \surname{Simon}}%
\affiliation{%
    Department of Biostatistics,%
    University of Washington, Seattle,%
    Washington 98195--1560, USA
}

\author{Ariel \surname{Rokem}}%
\affiliation{%
    Department of Psychology,%
    University of Washington, Seattle,%
    Washington 98195--1560, USA
}

\date{\today}

% Please keep the abstract below 300 words
\input{output}
\input{abstract}

\maketitle

\input{sections/introduction}
\input{sections/results}
\input{sections/conclusion}
\input{sections/methods}

\begin{acknowledgments}
    {\it Acknowledgments.-}
    This work was supported by BRAIN Iniative grant 1RF1MH121868-01 from the
    National Institutes for Mental Health and by a grant from the Gordon \& Betty
    Moore Foundation and the Alfred P. Sloan Foundation to the University of
    Washington eScience Institute Data Science Environment. We would like to thank
    Scott Murray for a useful discussion of the SGL method and Mareike Grotheer for
    helpful comments on the manuscript.
\end{acknowledgments}

\printfigures

\bibliographystyle{naturemag}
\bibliography{paper}

### 2021-01-21 reports/manuscript/sections/__latexindent_temp.tex
ENH: update paper and figures

\section*{Introduction}

>>>>

\subsection*{Introduction}

\begin{figure}[!h]
    \vspace{0.25cm}
    \rlap{
        \hspace{0.5em}
        \includegraphics[width=0.98\columnwidth]{methods-brains.pdf}
    }
    \vspace{-9.75cm}
    \includegraphics[width=\columnwidth]{method-quad.pdf}
    {\phantomsubcaption\label{fig:methods:tractogram}}
    {\phantomsubcaption\label{fig:methods:cst}}
    {\phantomsubcaption\label{fig:methods:tract-profile}}
    {\phantomsubcaption\label{fig:methods:group-structure}}
    \caption{{\bf Tractometry data lineage}
        \label{fig:methods}
        {\bf (a)} Whole brain tractography generates streamlines approximating
        the trajectories of white matter connections.
        {\bf (b)} Tractometry classifies these streamlines into anatomical bundles.
        In this case, we show the left corticospinal tract (CST\_L) over a mid-saggital
        anatomical slice.
        {\bf (c)} Tractometry further extracts bundle profiles,
        quantifications of various diffusion metrics along the length of the
        fiber bundle. Here, we show one subject's fractional anisotropy (fa)
        profile for CST\_L.
        {\bf (d)} the phenotypical target data and tractometric features can
        be organized into a linear model, $\hat{y} = \mathbf{X}
        \hat{\beta}$. The feature matrix $\mathbf{X}$ is color-coded
        to reveal a natural group structure: the left (orange) group
        contains $k$ features from the CST\_L, the middle (green) group
        contains $k$ features from the left cingulum cingulate (CGC\_L),
        and the right (blue) group
        contains $k$ features from the left arcuate (ARC\_L). The coefficients in
        $\hat{\beta}$ follow the same natural grouping.
        % Cross-validation:
        % we evaluate model quality using a nested $k$-fold cross validation
        % scheme. At level-0, the input data is decomposed into $k$
        % shuffled groups and optimal hyperparameters are found for the level-0
        % training set.
        % To avoid overfitting, the optimal hyperparameters are themselves
        % evaluated using a cross-validation scheme taking place
        % at level-1 of the decomposition, where each
        % level-0 training set is further decomposed into five shuffled
        % groups. For the ALS and WH data, $k=10$, while for the HBN and
        % Cam-CAN data, $k=5$.
    }
\end{figure}

\todo[inline]{The introduction section should be shortened for Nature methods
I think.}

### 2021-01-21 reports/manuscript/sections/conclusion.tex
ENH: update paper and figures

\section*{Conclusion}

@@

Taken together, these results demonstrate the promise of the group-regularized
regression approach. Even at the scale of dozens of subjects, the results
provided by SGL are both accurate, as well as interpretable
\cite{Murdoch2019-ax}: tractometry capitalizes on domain knowledge to engineer
meaningful features; SGL scores these features based on their relative
importance; enables a visualization of these feature importance scores in the
anatomical coordinate frame of the bundles (e.g., Figures \ref{fig:class-results}
and fig:regress-beta) and provides a means to understand model errors (e.g.,
Figure \ref{fig:class-errors}) . Thus, this multivariate analysis approach both
(a) achieves high cross-validated accuracy for precision medicine applications
of dMRI data and (b) identifies relevant features of brain anatomy that can
further our neuroscientific understanding of clinical disorders.

>>>>

\subsection*{Discussion}

@@

One drawback of our approach is also evident in the age regression results.
There are portions of the bundle profiles in
\cref{fig:age-results:wh-profile,fig:age-results:hbn-profile,fig:age-results:cc-profile}
for which the $\hat{\beta}$ coefficients are zero but for which the bundle
profiles clearly contain an age differentiating signal. While SGL correctly
identifies certain features as important, it is not guaranteed to identify
\emph{all} of the important features. It will identify a parsimonious set of
important features, dense enough to predict the phenotype and sparse enough
(at either the group or global level) to satisfy the sparsity constraints.

Taken together, these results demonstrate the promise of the
group-regularized regression approach. Even at the scale of dozens of
subjects, the results provided by SGL are both accurate, as well as
interpretable \cite{Murdoch2019-ax}: tractometry capitalizes on domain
knowledge to engineer meaningful features; SGL scores these features based on
their relative importance; enables a visualization of these feature
importance scores in the anatomical coordinate frame of the bundles (e.g.,
\cref{fig:class-results:coefs3d,fig:age-results:wh-coefs3d,fig:age-results:hbn-coefs3d,fig:age-results:cc-coefs3d})
and provides a means to understand model errors (e.g.,
\cref{fig:class-results:false-positives,fig:class-results:false-negatives}) .
Thus, this multivariate analysis approach both (a) achieves high
cross-validated accuracy for precision medicine applications of dMRI data and
(b) identifies relevant features of brain anatomy that can further our
neuroscientific understanding of clinical disorders.

### 2021-01-21 reports/manuscript/sections/methods.tex
Add zenodo DOI
ENH: update paper and figures

\section*{Materials and methods}

\subsection*{Data}

Two different previously-published datasets were used here:

\begin{enumerate}

\item Diffusion MRI from a previous study of the corticospinal
tract (CST) in patients with amyotrophic lateral sclerosis
(ALS \cite{sarica2017corticospinal}), containing data from 24 ALS
patients and 24 demographically matched healthy controls. These data
were measured in a GE Discovery 750 3T MRI scanner at the Institute
of Bioimaging and Molecular Physiology in Catanzaro. Informed consent
was provided as approved by the Ethical Committee of the University
``Magna Graecia'' of Catanzaro. Voxel resolution was \num{2x2x2}
$\text{mm}^3$ and 27 non-colinear directions were measured with a
$b=1000$ $\frac{\text{sec}}{\text{mm}^2}$. Here, both motion correction
and eddy current correction were applied before modeling the diffusion
tensor in each voxel. We will refer to this dataset as ALS.

\item Diffusion MRI data from a previous study of properties of
the white matter across the lifespan \cite{yeatman2014lifespan},
containing dMRI data from 76 subjects with ages 6-50. These data were
measured in a GE Discovery 750 3T MRI scanner at the Stanford Center
for Cognitive and Neurobiological Imaging. The Stanford University
IRB approved the procedures of this study. Informed consent was
obtained from each adult participant, and assent for participation
was provided by parents/guardians for children. Voxel resolution was
\num{2x2x2}$\text{mm}^3$ with 96 non-colinear directions measured with a
$b=2000$ $\frac{\text{sec}}{\text{mm}^2}$ and 30 non-colinear directions
measured with a $b=1000$ $\frac{\text{sec}}{\text{mm}^2}$. Data was
preprocessed to correct for subject motion and the diffusion tensor
model \cite{basser1994mr} was fit in every voxel, using a robust fit
\cite{chang2005restore}. These data were acquired using a dual spin echo
sequence, in which there is sufficient time for eddy currents to subside
between the application of the gradients and the image acquisition, so
no eddy current correction was applied. We will refer to this dataset
as WH.

\end{enumerate}

Data from both of these studies was processed in a similar manner,
using the Matlab Automated Fiber Quantification toolbox (AFQ)
\cite{yeatman2012tract}: streamlines representing fascicles of white
matter tracts were generated using a determinstic tractography algorithm
that follows the prinicpal diffusion direction of the diffusion tensor
in each voxel (STT) \cite{basser2000vivo}. Major tracts were identified
using multiple criteria: streamlines are selected as candidates for
inclusion in a bundle of streamlines that represents a tract if they
pass through known inclusion ROIs and do not pass through exclusion
ROIs \cite{wakana2007reproducibility}. In addition, a probabilistic
atlas is used to exclude streamlines which are unlikely to be part of
a tract \cite{Hua2008-sh}. Each streamline is resampled to 100 nodes
and the robust mean at each location is calculated by estimating the 3D
covariance of the location of each node and excluding streamlines that
are more than 5 standard deviations from the mean location in any node.
Finally, a bundle profile of tissue properties in each bundle was created
by interpolating the value of MRI maps of these tissue properties to the
location of the nodes of the resampled streamlines designated to each
bundle. In each of 100 nodes, the values are summed across streamlines,
weighting the contribution of each streamline by the inverse of the
mahalnobis distance of the node from the average of that node across
streamlines. This means that streamlines that are more representative of
the tract contribute more to the bundle profile, relative to streamlines
that are on the edge of the tract.

This process creates bundle profiles, in which diffusion measures
are quantified and averaged along twenty major fiber tracts. Here,
we use only the mean diffusivity (MD) and the fractional anisotropy
(FA) of the diffusion tensor, but additional dMRI-based maps or maps
based on other quantitative MRI measurements can also be used. These
bundle profiles, along with the phenotypical data we wish to explain
or predict, form the input to the SGL algorithm. In a domain-agnostic
machine learning context, the phenotypical data comprise the target
variables while the bundle profiles form the feature or predictor
variables (See Fig~\ref{fig:group-structure}).

\subsection*{Sparse Group Lasso}

Before fitting a model to the data, imputation and standardization are
performed. Missing node values (e.g., in cases where AFQ designates a node as
not-a-number) are imputed via linear interpolation. Care is taken not to
interpolate across the boundaries between different bundles. Some diffusion
metrics will have naturally larger variance than others and may therefore
dominate the objective function and make the SGL estimator unable to learn from
the lower variance metrics. For example, fractional anisotropy (FA) is bounded
between zero and one and could be overwhelmed by an unscaled higher-variance
metric like the mean diffusivity (MD). To prevent this we remove each feature's
mean and scale it to unit variance (z-score) using the
\lstinline{StandardScaler} from scikit-learn \cite{scikit-learn}. Scaling is
performed separately within each cross-validation set's training or testing data
to prevent leakage of information between the testing and training
sets\cite{kaufman2012leakage}.

After scaling and imputation, the tractometry data and target
phenotypical data can be organized in a linear model:
\begin{equation}
    y = \mathbf{X} \beta + \epsilon,
    \label{eq:lm}
\end{equation}
where $y$ is the phenotype -- categorical, such as a clinical diagnosis,
or continuous numerical, such as the subject's age. The tractometry
data is represented in the feature matrix $\mathbf{X}$, with rows
corresponding to different subjects, and columns corresponding
to diffusion measures at different nodes within each bundle. The
relationship between tractometric features and the phenotypic target is
characterized by the coefficients in $\beta$. The error term, $\epsilon$
is an unobserved random variable that captures the error in the model.
We denote our prediction of the targer phenotype as $\hat{y}$ and the
coefficients that produce this prediction as $\hat{\beta}$, so that
\begin{equation}
    \hat{y} = \mathbf{X} \hat{\beta},
    \label{eq:lm-approx}
\end{equation}
without the error term, $\epsilon$. In general, the feature matrix
$\mathbf{X}$ has dimensions $S \times (B \times N \times M)$, where $S$
is the number of subjects, $B$ is the number of white matter bundles,
$N$ is the number of nodes in each bundle, and $M$ is the number of
diffusion metrics calculated at each node. Typically, $B = 20$, $N =
100$, and $2 \le M \le 8$, resulting in $\sim 4,000 - 16,000$ features.
Conversely, many dMRI studies have between a few dozen and a few
hundred subjects, yielding a feature matrix that is wide and short.
Even in cases where more than a thousand subjects are measured (e.g.,
in the Human Connectome Project, where 1,200 subjects were measured
\cite{VanEssen2012}), the problem is ill-posed: the high dimensionality
of this data requires regularization to avoid overfitting and generate
interpretable results.

Here, we propose that in addition to regularizing the coefficients
in $\hat{\beta}$, we can also capitalize on our knowledge of the
group structure of the bundle profile features in $\mathbf{X}$. The
bundle-metric combinations form a natural grouping. For example, we
expect that MD features within the left arcuate fasciculus will
co-vary across individuals. Likewise for FA values within the right
corticospinal tract (CST) and so on. This group structure is represented
in Fig~\ref{fig:group-structure}, which depicts the linear model
$\hat{y} = \mathbf{X} \hat{\beta}$. Thus, we seek a regularization
approach that will fit a linear model with anatomically-grouped
covariates, where we expect to observe both groupwise sparsity, where
the number of groups (bundle/metric combinations) with at least one
non-zero coefficients is small, as well as within-group sparsity, where
the number of non-zero coefficients within each non-zero group is small.
The sparse group lasso (SGL) is a penalized regression technique that
satisfies exactly these criteria\cite{simon2013sparse}. It solves for a
coefficient vector
$\hat{\beta}$ that satisfies
\begin{equation}
    \hat{\beta} = \min_\beta \frac{1}{2}
    ||y - \displaystyle \sum_{\ell = 1}^{G}
    \mathbf{X}^{(\ell)} \beta^{(\ell)}||_2^2
    + \lambda_1 \displaystyle \sum_{\ell = 1}^{G}
    \sqrt{p_\ell} ||\beta^{(\ell)}||_2
    + \lambda_2 ||\beta||_1,
    \label{eq:sgl}
\end{equation}
where $G$ is the number of groups $\mathbf{X}^{(\ell)}$ is the submatrix
of $\mathbf{X}$ corresponding to group $\ell$, $\beta^{(\ell)}$ is
the coefficient vector for group $\ell$ and $p_\ell$ is the length of
$\beta^{(\ell)}$. In the tractomtetry setting, $G = T \times M$ and
$\forall \ell: p_\ell = 100$. The first term is the mean square error
loss, $L_{\text{mse}}$, as in the standard linear regression framework.
The second and third terms encourage groupwise sparsity and overall
sparsity, respectively. If $\lambda_1 = 0$ and $\lambda_2 = 1$, the
SGL reduces to the traditional lasso\cite{tibshirani1996regression}.
Conversely, if $\lambda_1 = 1$ and $\lambda_2 = 0$, the SGL reduces to
the group lasso\cite{yuan2006model}.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.65\textwidth]{dMRI_group_structure.png}
    \caption{{\bf dMRI group structure.}
        The phenotypical target data and tractometric features can
        be organized into a linear model, $\hat{y} = \mathbf{X}
        \hat{\beta}$. The feature matrix $\mathbf{X}$ is color-coded
        to reveal a natural group structure: the left (orange) group
        contains $k$ features from the inferior fronto-occipital
        fascicle (IFOF), the middle (green) group contains $k$ features
        from the corpus callosum, and the right (blue) group
        contains $k$ features from the uncinate. The coefficients in
        $\hat{\beta}$ follow the same natural grouping. Fascicle image
        reproduced with permission from Ref~\cite{yeatman2012tract}
        Figure 1.
    }
    \label{fig:group-structure}
\end{figure}

\subsubsection*{Incorporating transformations on $y$}

Often, the target variable $y$ is not in the domain in which the linear
model can be best fit to it. Equation \eqref{eq:lm-approx} can be slightly
modified as:
\begin{equation}
    \hat{y} = f^-1(\mathbf{X} \hat{\beta}),
    \label{eq:lm-transform}
\end{equation}
where the transformation function $f^{-1}$ characterizes the transform
applied to the data before fitting the linear coefficients. For example,
an often-used transform is the logarithmic transform:
\begin{equation}
    f(\hat{y}) = \log_n(\hat{y})
    \label{eq:log-nonlinearity}
\end{equation}
In this case, the model is parametrized by one additional fit parameter,
$n$.

\subsubsection*{Classification of categorical $y$}
When the phenotypical target variable is categorical, as in the case of
explaining or predicting the presence of a clinical diagnosis, the SGL is
readily adapted to logistic regression, where the probability of a target
variable belonging to an arbitrary defined ``true'' class is the logistic
function of the result of the linear model,
\begin{equation}
    p(\hat{y} = 1) = \frac{1}{1 + \exp(\mathbf{X} \hat{\beta})},
    \label{eq:logit}
\end{equation}
or equivalently, the mean squared error loss function in Eq~\eqref{eq:sgl} is
replaced with the cross-entropy loss, which for binary classification is the
negative log likelihood of the SGL classifier giving the ``true'' label:
\begin{equation}
    L_{\text{mse}} \rightarrow L_{\log} =
    -\left(y \log(p) + (1 - y) \log(1 - p)\right).
    \label{eq:logloss}
\end{equation}

\subsection*{Implementation, cross-validation and metaparameter optimization}

For given values of $\lambda_1$ and $\lambda_2$, the cost function in Eq~\eqref{eq:sgl} can be optimized using proximal gradient descent methods
\cite{parikh2014proximal} here implemented as a custom proximal operator that is
then optimized using the C-OPT library\cite{copt}. This supplies an estimate of
the optimal $\hat{\beta}$ given a particular set of values for the
meta-parameters $\lambda_1$ and $\lambda_2$.

To objectively evaluate the model and guard against over-fitting,
we used a nested cross-validation scheme, depicted in
Fig~\ref{fig:cross-val} for the categorical classification case.
The subjects (i.e. rows of the feature matrix $\mathbf{X}$ in
Fig~\ref{fig:group-structure} and Eq~\eqref{eq:lm}) are shuffled and
then decomposed into $k$ batches, hereafter referred to as folds. For
the ALS dataset we used $k=10$ and for the WH dataset $k=5$. For each
unique fold, we hold that fold out as the test\textsubscript{outer} set
and let the remaining data comprise the train\textsubscript{outer} set,
with the subscript indicating the depth of the nested decomposition.
We further decompose each train\textsubscript{outer} set into three
folds, and again for each unique fold, we hold out that fold as the
test\textsubscript{inner} set and let the remaining data comprise the
train\textsubscript{inner} set. At level-1 of the decomposition, we fit
an SGL model using fixed regularization meta-parameters $\lambda_1$
and $\lambda_2$, training the model using train\textsubscript{inner}
and evaluating the fit on test\textsubscript{inner}. We find
the optimal values for $\lambda_1$ and $\lambda_2$ using
hyperoptimization, implemented using the hyperopt library's \verb|fmin|
function\cite{Bergstra_2015} with a tree of Parzen estimators search
algorithm\cite{bergstra2011algorithms}. For continuous numerical $y$,
\verb|fmin| searches for meta-parameter values that minimize the median
absolute error. This can also be done minimizing the root of the mean
squared error (RMSE) or to maximizing the coefficient of determination
($R^2$). For binary categorical $y$ \verb|fmin| seeks to maximize the
classification accuracy. This can also be done maximizing the area
under the receiver operating curve (ROC AUC) or the average precision.
Using hyperoptimization, we find optimal regularization parameters and
$\hat{\beta}$ for each train\textsubscript{outer} set and then use those
to predict values for data in test\textsubscript{outer}. Thus each
subject in the dataset has a predicted phenotype derived from a model
that never saw its particular subject's data.

The above procedure describes $k$-fold cross validation. In fact, we use
repeated $k$-fold cross validation on the outer level of the decomposition, so
that the input data is decomposed into $k$ folds, three times. Thus, each
subject has three predicted phenotypes. We then take the mean predicted value to
summarize the performance of the model. In the classification case, the
shuffling into folds is stratified such that each fold has a population that
preserves the percentage of each class found in the larger input data.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.55\textwidth]{nested-cross-validation.pdf}
    \caption{{\bf Nested $k$-fold cross-validation scheme.}
        We evaluate model quality using a nested $k$-fold cross
        validation scheme. At level-0, the input data is decomposed
        three times into $k$ shuffled groups and optimal hyperparameters
        are found for the level-0 training set. Optimization of these
        hyperparameters requires the use of the hyperopt library and
        many repeated evaluations of an SGL model over a search space
        of possible regularization parameters. These evaluations take
        place at level-1 of the decomposition, where the level-0 training
        set is further decomposed into three shuffled groups.
        For the ALS data, $k=10$. For the WH data, $k=5$.
    }
    \label{fig:cross-val}
\end{figure}

\subsection*{Software implementation}

The full software implementation of the SGL approach presented here is available
in a Python software package called AFQ-Insight, which is developed publicly in
\url{https://github.com/richford/afq-insight}. The version of the code used to
produce the results herein is also available in \ariel{Need to add Zenodo DOI}.
AFQ-Insight reads the target and feature data that has been processed by AFQ
from comma separated value (CSV) files conforming to the AFQ-Browser data
format\cite{yeatman2018browser} and represents them internally as
\lstinline{DataFrame} objects from the pandas Python
library\cite{mckinney2010data}. The software provides different options for
imputing missing data in the feature matrix. Missing interior nodes are imputed
using linear interpolation. For missing exterior nodes, the user may choose
between linear extrapolation and constant forward(back)-fill. Imputation uses
only values from adjacent nodes in the same white matter bundle in the same
subject so there is no danger of data leakage from other subjects. It uses the
scikit-learn\cite{scikit-learn} library to decompose input data into separate
test and train datasets, to scale each feature to have zero mean and
unit variance, and to evaluate each fit in the hyperparameter search using
appropriate classification and regression metrics such as accuracy, area
under the receiver operating curve (AUC ROC), and coefficient of determination
($R^2$). For each set of hyperparameters, we solve the SGL using a custom
proximal operator supplied to the C-OPT library\cite{copt}. Appropriate
hyperparameters are found using the hyperopt library\cite{Bergstra_2015}.

>>>>

\subsection*{Methods}

Methods and associated references are available in \cref{supp:sec:methods} of
the supplemental material \cite{supplement}.
